\chapter{Linear Geometry}
\label{chap:linear-geometry}

This chapter is dedicated to fostering a geometric intuition about
\hyperref[def:vector]{vectors} as introduced in
\myref{section}{sec:describing-solution-sets-of-linear-systems}. There, we used
them as a convenient way of grouping data. They, however, are also apt
representations of a geometric concept -- the concept of a `shift in space'.

Before we move on, though, we need elucidate what we mean by `geometry' and
`space'. The former word has a widespread connotation of `drawing stuff using a
ruler and a compass on a sheet of paper'. This is hardly the \emph{geometry} we
have in mind in this text. Although circles, lines and triangles are geometric
objects to us as well, the fundamental notion of geometry, which is scarcely
properly discussed in high-school setting, is \emph{space}.

\emph{Space} is the `place where we do geometry', in essence. We begin by
arguing the most intuitive way of describing space is that of a set of
dimensions (or directions of movement) where each dimension is a
\emph{continuum}, meaning, in whichever direction I move, there are no holes
along the way. We trust the first \emph{continuum} (a `set without holes') dear
readers encountered, are the real numbers, $\R$. They seem a good candidate for
the definition of space.

Firstly, we define the \emph{one-dimensional} space to be exactly $\R$. You
might wonder how this makes any sense. Well, a space with just one dimension is
an infinite line like below.
\begin{center}
 \begin{tikzpicture}
  \draw[arrows={Latex[width=6pt,length=8pt]-Latex[width=6pt,length=8pt]}]
   (0,0) to (10,0);
 \end{tikzpicture}
\end{center}
How is this related to $\R$? Quite trivially. Pick any point on the line and
label it $0$. Then pick yet another different point and label it $1$. Like so.
\begin{center}
 \begin{tikzpicture}
  \draw[arrows={Latex[width=6pt,length=8pt]-Latex[width=6pt,length=8pt]}]
   (0,0) to (10,0);
  \draw[thick,BrickRed] (5,0.1) -- (5,-0.1);
  \node[BrickRed] at (5,-0.4) {$0$};
  \draw[thick,RoyalBlue] (6,0.1) -- (6,-0.1);
  \node[RoyalBlue] at (6,-0.4) {$1$};
 \end{tikzpicture}
\end{center}
Just like that, we have forged a \emph{correspondence} between an infinite line
and the real numbers. Any real number $r \in \R$ corresponds to the point on the
line distant exactly $|r|$ from $0$ -- to the right if it's positive, and to the
left if it's negative. For example, the number $\clg{-3.14}$ is represented as
the following \clg{point}.
\begin{center}
 \begin{tikzpicture}
  \draw[arrows={Latex[width=6pt,length=8pt]-Latex[width=6pt,length=8pt]}]
   (0,0) to (10,0);
  \draw[thick,BrickRed] (5,0.1) -- (5,-0.1);
  \node[BrickRed] at (5,-0.4) {$0$};
  \draw[thick,RoyalBlue] (6,0.1) -- (6,-0.1);
  \node[RoyalBlue] at (6,-0.4) {$1$};
  \node[ForestGreen] at (1.86,-0.4) {$-3.14$};
  \draw[thick,ForestGreen] (1.86,0.1) -- (1.86,-0.1);
  \draw[decorate,decoration={brace,raise=5pt,amplitude=8pt}] (1.86,0) -- (5,0);
  \node at (3.43,0.7) {$3.14$};
 \end{tikzpicture}
\end{center}

Why stop at one dimension? We may add as many dimensions as we like by simply
inventing new directions of movement. Knowing that one direction of movement
amounts to the set of real numbers, we get as many directions as we choose to
include copies of the set of real numbers.

To make the preceding paragraph more tangible, let's first add a second
dimension, which we typically depict as another line perpendicular to the first
one and meeting it at $0$. Like this.
\begin{center}
 \begin{tikzpicture}
  \draw[arrows={Latex[width=4pt,length=6pt]-Latex[width=4pt,length=6pt]}]
   (0,0) to (4,0);
  \draw[arrows={Latex[width=4pt,length=6pt]-Latex[width=4pt,length=6pt]}]
   (2,2) to (2,-2);
  \draw[thick,BrickRed] (2,0.1) -- (2,-0.1);
  \draw[thick,BrickRed] (2.1,0) -- (1.9,0);
  \node[BrickRed] at (1.7,-0.3) {$0$};
  \draw[thick,RoyalBlue] (3,0.1) -- (3,-0.1);
  \node[RoyalBlue] at (3,-0.3) {$1$};
  \draw[thick,RoyalBlue] (1.9,1) -- (2.1,1);
  \node[RoyalBlue] at (1.7,1) {$1$};
 \end{tikzpicture}
\end{center}
Now, we have the entire \emph{plane} to move about. We are free to tread as far
right (or left) as we wish and as far up (or down) as we wish. Therefore, each
point here is given two numbers -- one on the axis of horizontal movement and
one on the axis of vertical movement.
\begin{center}
 \begin{tikzpicture}
  \draw[arrows={Latex[width=4pt,length=6pt]-Latex[width=4pt,length=6pt]}]
   (0,0) to (4,0);
  \draw[arrows={Latex[width=4pt,length=6pt]-Latex[width=4pt,length=6pt]}]
   (2,2) to (2,-2);
  \draw[thick,BrickRed] (2,0.1) -- (2,-0.1);
  \draw[thick,BrickRed] (2.1,0) -- (1.9,0);
  \node[BrickRed] at (1.7,-0.3) {$0$};
  \draw[thick,RoyalBlue] (3,0.1) -- (3,-0.1);
  \node[RoyalBlue] at (3,-0.3) {$1$};
  \draw[thick,RoyalBlue] (1.9,1) -- (2.1,1);
  \node[RoyalBlue] at (1.7,1) {$1$};
  \draw[thick,RoyalBlue] (1.9,1) -- (2.1,1);
  \node[ForestGreen] at (1.55,0.6) {$0.6$};
  \draw[thick,ForestGreen] (1.9,0.6) -- (2.1,0.6);
  \draw[thick,ForestGreen] (3.5,0.1) -- (3.5,-0.1);
  \node[ForestGreen] at (3.5,-0.3) {$1.5$};
  
  \draw[ForestGreen,dashed] (3.5,0) -- (3.5,0.6);
  \draw[ForestGreen,dashed] (2,0.6) -- (3.5,0.6);
  \node[ForestGreen,circle,fill,inner sep=1.5pt] at (3.5,0.6) {};
  \node[ForestGreen] at (4.3,0.9) {$(1.5,0.6)$};
 \end{tikzpicture}
\end{center}
We call the set of all ordered pairs of real numbers the \emph{cartesian
product} of real numbers with themselves and denote it $\R \times \R$ (or $\R^2$
for short). Hence, the two-dimensional space (also called \emph{plane}) is just
equivalent to $\R^2$.

We needn't stop at two dimensions. As stated before, we may add as many
directions of movement as desired. The preceding small examples justify the
following definition.

\begin{definition}{Space}{space}
 We call the set $\R^{n}$ of all (ordered) $n$-tuples of real numbers the
 \emph{$n$-dimensional (real) space}. 
\end{definition}

We claimed that a \emph{vector} represents a shift in space. We now make that
idea precise. Consider the points $(a_1,a_2)$ and $(b_1,b_2)$ in $\R^2$. There
exists a straight path from one to the other. The trajectory of such path can be
represented as a pair of numbers where the first one signifies the distance we
are to travel horizontally to reach $b_1$ from $a_1$ and the second number the
vertical distance from $a_2$ to $b_2$. Clearly, the first distance equals $b_1 -
a_1$ and the second $b_2 - a_2$. We can collect these numbers into the vector
\[
 \mathbf{v} = 
 \begin{pmatrix}
  b_1 - a_1\\
  b_2 - a_2
 \end{pmatrix}
\]
which now serves as a representation of the \emph{movement} or \emph{shift} from
the point $(a_1,a_2)$ to the point $(b_1,b_2$). In this scenario, we call
$(a_1,a_2)$ the \emph{root} or \emph{source} of the vector $\mathbf{v}$ and the
point $(b_1,b_2)$ its \emph{end} or \emph{target}.

Naturally, this idea is easily scaled to higher dimensions. Given points
\begin{align*}
 a &= (a_1,a_2,\ldots,a_n),\\
 b &= (b_1,b_2,\ldots,b_n)
\end{align*}
in $\R^{n}$, the vector with \emph{source} $a$ and \emph{target} $b$ is exactly
\[
 \begin{pmatrix}
  b_1 - a_1\\
  b_2 - a_2\\
  \vdots\\
  b_n - a_n
 \end{pmatrix}.
\]
\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \tkzInit[xmin=-1,xmax=4,ymin=-1,ymax=4]
  \tkzDrawX[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDrawY[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDefPoints{1/2/a,3/3/b,1/0/a1,0/2/a2,3/0/b1,0/3/b2}
  \tkzDrawPoints[size=4,fill=black](a1,a2,b1,b2)
  \tkzLabelPoint[below=1mm](a1){$1$}
  \tkzLabelPoint[below=1mm](b1){$3$}
  \tkzLabelPoint[left=1mm](a2){$2$}
  \tkzLabelPoint[left=1mm](b2){$3$}

  \tkzDrawSegment[-Latex,thick,Fuchsia,shorten <=4pt,shorten >=4pt](a,b)
  \tkzLabelSegment[pos=0.5,below,Fuchsia](a,b){$\mathbf{v}$}
  \tkzDrawSegments[dashed](a1,a a2,a b1,b b2,b)
  \tkzDrawPoints[size=6,BrickRed,fill=BrickRed](a,b)
 \end{tikzpicture}
 \caption{Depiction of a vector $\clm{\mathbf{v}}$ with source $\clr{(1,2)}$ and
 target $\clr{(3,3)}$.}
 \label{fig:vector}
\end{figure}

Observe how this idea of vector being a \emph{shift} sings in unison with the
definitions of \hyperref[def:adding-vectors]{vector addition} and
\hyperref[def:multiplying-vector-by-a-number]{scalar multiplication}. The sum of
the vector of source $a$ and target $b$ with the vector of source $b$ and target
$c$ is the vector of source $a$ and target $c$ (cf.
\myref{figure}{fig:vector-addition}). This is testified by the simple
calculation
\[
 \begin{pmatrix}
  b_1 - a_1\\
  b_2 - a_2\\
  \vdots\\
  b_n - a_n
 \end{pmatrix}
 + 
 \begin{pmatrix}
  c_1 - b_1\\
  c_2 - b_2\\
  \vdots\\
  c_n - b_n
 \end{pmatrix}
 =
 \begin{pmatrix}
  b_1 - a_1 + c_1 - b_1\\
  b_2 - a_2 + c_2 - b_2\\
  \vdots\\
  b_n - a_n + c_n - b_n
 \end{pmatrix}
 = 
 \begin{pmatrix}
  c_1 - a_1\\
  c_2 - a_2\\
  \vdots\\
  c_n - a_n
 \end{pmatrix}
\]
for points $(a_1,\ldots,a_n), (b_1,\ldots,b_n)$ and $(c_1,\ldots,c_n)$ in
$\R^{n}$.

\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \tkzInit[xmin=-1,xmax=4,ymin=-1,ymax=4]
  \tkzDrawX[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDrawY[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDefPoints{1/2/a,3/3/b,1/0/a1,0/2/a2,3/0/b1,0/3/b2,4/1/c,4/0/c1,0/1/c2}
  \tkzDrawPoints[size=4,fill=black](a1,a2,b1,b2,c1,c2)
  \tkzLabelPoint[below=1mm](a1){$1$}
  \tkzLabelPoint[below=1mm](b1){$3$}
  \tkzLabelPoint[below=1mm](c1){$4$}
  \tkzLabelPoint[left=1mm](a2){$2$}
  \tkzLabelPoint[left=1mm](b2){$3$}
  \tkzLabelPoint[left=1mm](c2){$1$}

  \tkzDrawSegment[-Latex,thick,BrickRed,shorten <=4pt,shorten >=4pt](a,b)
  \tkzDrawSegment[-Latex,thick,RoyalBlue,shorten <=4pt,shorten >=4pt](b,c)
  \tkzDrawSegment[-Latex,thick,ForestGreen,shorten <=4pt,shorten >=4pt](a,c)
  \tkzLabelSegment[pos=0.4,above,BrickRed](a,b){$\mathbf{v}$}
  \tkzLabelSegment[pos=0.4,right,RoyalBlue](b,c){$\mathbf{w}$}
  \tkzLabelSegment[pos=0.4,below=1mm,ForestGreen](a,c){$\mathbf{v} + \mathbf{w}$}
  \tkzDrawSegments[dashed](a1,a a2,a b1,b b2,b c1,c c2,c)
  \tkzDrawPoints[size=6,fill=black](a,b,c)
 \end{tikzpicture}
 \caption{The sum $\clg{\mathbf{v} + \mathbf{w}}$ of vectors $\clr{\mathbf{v}}$
 and $\clb{\mathbf{w}}$.}
 \label{fig:vector-addition}
\end{figure}

Similarly, the concept of scalar multiplication is transmitted as enlarging or
shortening (possibly reversing if the scalar is negative) of the vector in
question (cf. \myref{figure}{fig:scalar-multiplication}).

\begin{figure}[ht]
 \centering
 \begin{subfigure}[b]{.45\textwidth}
  \centering
  \begin{tikzpicture}
   \tkzInit[xmin=-1,xmax=5,ymin=-1,ymax=4]
   \tkzDrawX[arrows={-Latex[width=4pt,length=6pt]},label=]
   \tkzDrawY[arrows={-Latex[width=4pt,length=6pt]},label=]
   \tkzDefPoints{1/2/a,3/3/b,1/0/a1,0/2/a2,3/0/b1,0/3/b2,5/4/c,5/0/c1,0/4/c2}
   \tkzDrawPoints[size=4,fill=black](a1,a2,b1,b2,c1,c2)
   \tkzLabelPoint[below=1mm](a1){$1$}
   \tkzLabelPoint[below=1mm](b1){$3$}
   \tkzLabelPoint[below=1mm](c1){$5$}
   \tkzLabelPoint[left=1mm](a2){$2$}
   \tkzLabelPoint[left=1mm](b2){$3$}
   \tkzLabelPoint[left=1mm](c2){$4$}

   \tkzDrawSegment[-Latex,BrickRed,thick,shorten <=4pt,shorten >=4pt](a,b)
   \tkzLabelSegment[pos=0.4,above,BrickRed](a,b){$\mathbf{v}$}
   \tkzDrawSegments[dashed](a1,a a2,a b1,b b2,b c1,c c2,c)
   \tkzDrawPoints[size=6,fill=black](a,b,c)
  \end{tikzpicture}
 \end{subfigure}
 \begin{subfigure}[b]{.45\textwidth}
  \centering
  \begin{tikzpicture}
   \tkzInit[xmin=-1,xmax=5,ymin=-1,ymax=4]
   \tkzDrawX[arrows={-Latex[width=4pt,length=6pt]},label=]
   \tkzDrawY[arrows={-Latex[width=4pt,length=6pt]},label=]
   \tkzDefPoints{1/2/a,3/3/b,1/0/a1,0/2/a2,3/0/b1,0/3/b2,5/4/c,5/0/c1,0/4/c2}
   \tkzDrawPoints[size=4,fill=black](a1,a2,b1,b2,c1,c2)
   \tkzLabelPoint[below=1mm](a1){$1$}
   \tkzLabelPoint[below=1mm](b1){$3$}
   \tkzLabelPoint[below=1mm](c1){$5$}
   \tkzLabelPoint[left=1mm](a2){$2$}
   \tkzLabelPoint[left=1mm](b2){$3$}
   \tkzLabelPoint[left=1mm](c2){$4$}

   \tkzDrawSegment[-Latex,thick,ForestGreen,shorten <=4pt,shorten >=4pt](a,c)
   \tkzLabelSegment[pos=0.4,below=1mm,ForestGreen](a,c){$2 \cdot \mathbf{v}$}
   \tkzDrawSegments[dashed](a1,a a2,a b1,b b2,b c1,c c2,c)
   \tkzDrawPoints[size=6,fill=black](a,b,c)
  \end{tikzpicture}
 \end{subfigure}
 \caption{The multiple $\clg{2 \cdot \mathbf{v}}$ of the vector
 $\clr{\mathbf{v}}$.}
 \label{fig:scalar-multiplication}
\end{figure}

There is a point to be made here. \myref{Figure}{fig:vector-addition} only shows
addition of vectors such that the first one ends where the second begins. You
may object that this aligns not with the general
\hyperref[def:adding-vectors]{definition} you gave earlier. There was no talk
about sources or targets. Fret not, as a simple observation alleviates the issue
-- a vector itself in fact carries no information whatsoever about either its
source or its target. For example, the vector
\[
 \begin{pmatrix}
  1\\
  2\\
  3
 \end{pmatrix}
\]
can have as its source the point $(0, 1, 2)$ and as its target $(1, 3, 7)$; it
can equally well be rooted at $(8, -2, 3)$ and end in $(9, 0, 7)$ or start at
just about any point $(x,y,z)$ as long as it ends in $(x + 1, y + 2, z + 3)$.

\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \tkzInit[xmin=-2,xmax=3,ymin=-1,ymax=3]
  \tkzDrawX[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDrawY[arrows={-Latex[width=4pt,length=6pt]},label=]
  \tkzDefPoints{1/0/a1,2/2/a2,-1/-1/b1,0/1/b2,-2/1/c1,-1/3/c2,2/-1/d1,3/1/d2}
  \tkzDrawSegments[-Latex,thick,BrickRed,shorten <=4pt,shorten >=4pt](a1,a2
  b1,b2 c1,c2 d1,d2)
 \end{tikzpicture}
 \caption{The vector $\clr{\begin{psmallmatrix} 1 \\ 2 \end{psmallmatrix}}$
 drawn with four different sources and targets.}
 \label{fig:one-vector-four-positions}
\end{figure}

It follows that we have rightly interpreted \hyperref[def:adding-vectors]{the
sum} of $\mathbf{v}$ and $\mathbf{w}$ as `moving along $\mathbf{v}$ and then
continuing along $\mathbf{w}$' since we can always move the start of
$\mathbf{w}$ to the end of $\mathbf{v}$. This interpretation also hopefully
drives home the idea that a vector is simply a description of a \emph{shift} or
\emph{displacement} in space, not exactly a \emph{segment} or a \emph{path}. We
tend to regard the latter two as having a fixed start and end which, as we just
saw, vectors don't do.

There is one formal consequence of this idea. Since vectors are free to be
rooted anywhere, why don't we simply make our lives easy and root them at the
origin -- the point $(0,\ldots,0)$? Doing so would make vectors formally
indistinguishable from points in space. Indeed, we may now trivially draw a
relationship between a point $(a_1,\ldots,a_n) \in \R^{n}$ and the vector
\[
 \begin{pmatrix}
  a_1\\
  a_2\\
  \vdots\\
  a_n
 \end{pmatrix}
\]
by consensus rooted at $(0,\ldots,0)$. Intuitively, the former is a specific
place in $n$-dimensional space and the latter is a shift from the origin to that
place. Formally, however, there is no difference at all to be found. That said,
we shall henceforth regard $\R^{n}$ as the set of all points with $n$ real
coordinates as well as the set of all vectors with $n$ real entries, whichever
one is more convenient in a given context. Expressed symbolically, we may write
\[
 \R^{n} = \{(a_1,\ldots,a_n) \mid a_1,\ldots,a_n \in \R\} = \left\{
  \begin{pmatrix}
   a_1\\
   a_2\\
   \vdots\\
   a_n
  \end{pmatrix} \mid a_1,\ldots,a_n \in \R
 \right\}.
\]

The concept of a \emph{shift} typically brings with itself two important
characteristics -- \emph{length} and \emph{direction}. In natural language, we
habitually express shifts by saying `Make three steps forward.' or `Turn
slightly right and then keep driving for 4 kilometres.' We should ask: `What is
the \emph{length} and \emph{direction} of a vector?' The rest of this chapter
lays down an answer to this question.

\input{linear-geometry/length}
\input{linear-geometry/angle}
