\subsection{Three-dimensional Linear Systems}
\label{ssec:three-dimensional-linear-systems}

Stepping up the game a little, we're taking a look at linear systems in three
variables. Just as a linear equations in two variables are lines in the real
plane, linear equations in three variables depict geometric objects called
`planes' in the three-dimensional real space, $\R^3$. They form the last class
of linear systems that can be efficiently visualized; with linear systems in
more variables being generally out of our perceptive reach.

Planes are the \emph{straight} objects in three-dimensional kind of sense. They
lock one direction of movement by making one variable wholly dependent on the
other two. An illustration is provided in \myref{figure}{fig:plane}.

\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \begin{axis}[
   view={60}{10},
   xlabel={$x$},
   ylabel={$y$},
   zlabel={$z$},
   zlabel style={rotate=90},
   xtick=\empty,
   ytick=\empty,
   ztick=\empty,
   xmin=-3,
   xmax=3,
   ymin=-5,
   ymax=5
  ]
   \addplot3[
    surf,
    colormap/hot,
    shader=faceted interp,
    ]({x}, {y}, {(2*x - y + 3) / 3});
  \end{axis}
 \end{tikzpicture}

 \caption{A \clb{plane} defined by the equation $2x - y - 3z = -3$.}
 \label{fig:plane}
\end{figure}

An \emph{underdetermined} system in three variables can contain either one or
two linear equations. In the former case, only one variable is dependent on the
other two -- we shall often call the independent variables by names such as
\emph{free variables} or \emph{parameters}. Both these names signify that a
substitution of any pair of real numbers in lieu of the two \emph{free
variables} yields a solution of the system.

For instance, the linear equation in \myref{figure}{fig:plane} is effectively a
linear system in three variables. We can choose any of the three variables to be
dependent and leave the other two free, giving thus three different descriptions
of \emph{the same} solution set. The following
equation~\eqref{eq:three-vars-one-eq} shows all of them with the chosen
dependent variable written on the left in typewriter font.
\begin{equation}
 \label{eq:three-vars-one-eq}
 \begin{array}{r l}
  \mathtt{x:} & \left\{ \left( \frac{y + 3z - 3}{2}, y, z \right) \mid y,z \in \R
  \right\}\\[1em]
  \mathtt{y:} & \left\{ \left( x, 2x - 3z + 3, z \right) \mid x,z \in \R
  \right\}\\[1em]
   \mathtt{z:} & \left\{ \left( x, y, \frac{2x - y + 3}{3} \right) \mid x,y \in
   \R \right\}
 \end{array}
\end{equation}

Linear systems in three variables and two equations are also underdetermined.
Geometrically, they correspond to arrangements of two planes in space. Those two
planes can either be parallel -- leading to the system having no solution -- or
not -- intersecting in a straight line describable as a set of triples with
exactly one free variable. In a case similar to two-dimensional linear systems,
putting the system in question into echelon form \emph{can} reveal (albeit not
always) its geometric nature.

Indeed, consider the system
\begin{equation}
 \label{eq:two-parallel-planes}
 \begin{array}{r c r c r c r}
  x & - & y & + & 2z & = & 2\\
  2x & - & 2y & + & 4z & = & 9
 \end{array}.
\end{equation}
Subtracting $\mathtt{II - 2 \cdot I}$ produces
\[
 \begin{array}{r c r c r c r}
  x & - & y & + & 2z & = & 2\\
    &   &   &   &  0 & = & 5
 \end{array},
\]
clearly a system with no solution. Subsequently, the two corresponding planes
are parallel to each other. See them depicted in
\myref{figure}{fig:two-parallel-planes}.

\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \begin{axis}[
   view={-60}{20},
   xlabel={$x$},
   ylabel={$y$},
   zlabel={$z$},
   zlabel style={rotate=90},
   xtick=\empty,
   ytick=\empty,
   ztick=\empty,
   xmin=-3,
   xmax=3,
   ymin=-5,
   ymax=5
  ]
   \addplot3[
    surf,
    colormap/hot,
    shader=faceted interp,
   ]({x}, {y}, {(2 - x + y) / 2});
   \addplot3[
    surf,
    colormap/viridis,
    shader=faceted interp,
    opacity=0.7,
   ]({x}, {y}, {(9 - 2*x + 2*y) / 4});
  \end{axis}
 \end{tikzpicture}

 \caption{The two parallel planes from the
 system~\eqref{eq:two-parallel-planes}.}
 \label{fig:two-parallel-planes}
\end{figure}

A system of two non-parallel planes is presented below.
\begin{equation}
 \label{eq:two-non-parallel-planes}
 \begin{array}{r c r c r c r}
  x & - & y & + & 2z & = & 2\\
  2x & + & 3y & - & z & = & -1
 \end{array}
\end{equation}
By subtracting, once again, $\mathtt{II - 2 \cdot I}$, we put into the following
echelon form.
\[
 \begin{array}{r c r c r c r}
  x & - & y & + & 2z & = & 2\\
    &   & 5y & - & 5z & = & -5
 \end{array}
\]
The algorithm of Gauss-Jordan elimination limits our choice of free variables to
the ones left in the last row. We are hence to set either $y$ or $z$ loose while
caging the latter. Custom dictates to label all variables but the first of the
last row as \emph{parameters}, making $z$ the victor. The rest is just
back-substitution. We calculate $y = z - 1$ and substitute into the first
equation to receive
\[
 x - (z - 1) + 2z = 2, \quad \text{hence} \quad x = 1 - z.
\]
It follows that \emph{one possible} description of the solution set of the
system~\eqref{eq:two-non-parallel-planes} is
\[
 \left\{ (1 - z, z - 1, z) \mid z \in \R\right\}.
\]
See it depicted in \myref{figure}{fig:two-non-parallel-planes}.
\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \begin{axis}[
   view={60}{10},
   xlabel={$x$},
   ylabel={$y$},
   zlabel={$z$},
   zlabel style={rotate=90},
   xtick=\empty,
   ytick=\empty,
   ztick=\empty,
   xmin=-3,
   xmax=3,
   ymin=-5,
   ymax=5
  ]
   \addplot3[
    surf,
    colormap/hot,
    shader=faceted interp,
   ]({x}, {y}, {2*x + 3*y + 1});
   \addplot3[
    surf,
    colormap/viridis,
    shader=faceted interp,
    opacity=0.6
   ]({x}, {y}, {(2 - x + y) / 2});
  \addplot3[
   color=BrickRed,
   thick,
   smooth
  ]({x},{-x},{1-x});
  \end{axis}
 \end{tikzpicture}

 \caption{The two non-parallel planes from the
 system~\eqref{eq:two-non-parallel-planes} and their \clr{intersection}.}
 \label{fig:two-non-parallel-planes}
\end{figure}

Reaching the apex of `ideal' linear systems in three variables and three
equations, we stop to ponder the number of arrangements of three planes in
three-dimensional space. There are two obvious ones:
\begin{enumerate}
 \item All three planes are parallel to each other.
 \item Only two planes are parallel to each other.
\end{enumerate}
Corresponding to (1), resp. (2), is a linear system with two equations, resp.
one equation, with no solution.

As an example, consider the system
\begin{equation}
 \label{eq:two-parallel-one-not}
 \begin{array}{r c r c r c r}
  -x & + & 2y & - & z & = & 4\\
  x & - & 6y & + & z & = & 1\\
  x & - & 2y & + & z & = & 3
 \end{array}.
\end{equation}
Its echelon form looks like this:
\[
 \begin{array}{r c r c r c r}
  -x & + & 2y & - & z & = & 4\\
     & & -4y & & & = & 5\\
     & & & & 0 & = & 7
 \end{array}.
\]
Clearly, the third equation has no solution while the second does. This
particular system's geometrical interpretation aligns with case (2) above. It is
shown in \myref{figure}{fig:two-parallel-one-not}.
\begin{figure}[ht]
 \centering
 \begin{tikzpicture}
  \begin{axis}[
   view={110}{-10},
   xlabel={$x$},
   ylabel={$y$},
   zlabel={$z$},
   zlabel style={rotate=90},
   xtick=\empty,
   ytick=\empty,
   ztick=\empty,
   xmin=-3,
   xmax=3,
   ymin=-5,
   ymax=5
  ]
   \addplot3[
    surf,
    colormap/hot,
    shader=faceted interp
   ]({x}, {y}, {-x + 6*y + 1});
   \addplot3[
    surf,
    colormap/viridis,
    shader=faceted interp,
    opacity=0.6
   ]({x}, {y}, {-x + 2*y - 4});
   \addplot3[
    surf,
    colormap/viridis,
    shader=faceted interp,
    opacity=0.6
   ]({x}, {y}, {-x + 2*y + 3});
   \addplot3[
    color=BrickRed,
    thick,
    smooth
   ]({x},{1/2},{4-x});
   \addplot3[
    color=BrickRed,
    thick,
    smooth
   ]({x},{-5/4},{-x - 13/2});
  \end{axis}
 \end{tikzpicture}

 \caption{Depiction of the system~\eqref{eq:two-parallel-one-not}.}
 \label{fig:two-parallel-one-not}
\end{figure}

Finally, there are three other possible arrangements of three planes in space:
\begin{enumerate}
 \setcounter{enumi}{2}
 \item non-parallel planes that fail to have a common intersection (the
  so-often-called `tent' configuration);
 \item non-parallel planes that meet in a single point;
 \item non-parallel planes that meet in a single line.
\end{enumerate}
Of course, as always, the echelon form of a linear system can be used to
distinctively label it as any of these cases. For example, the echelon form of
the system
\begin{equation}
 \label{eq:tent}
 \begin{array}{r c r c r c r}
  x & + & 2y & - & z & = & -1\\
  2x & - & 3y & + & 2z & = & 4\\
  -x & + & 5y & - & 3z & = & 10
 \end{array}
\end{equation}
can easily be computed to be
\[
 \begin{array}{r c r c r c r}
  x & + & 2y & - & z & = & -1\\
    & & -7y & + & 4z & = & 6\\
    & & & & 0 & = & 15
 \end{array}.
\]

