\section{Applications}
\label{sec:applications}

In this final section of \myref{chapter}{chap:linear-systems}, we focus on some
`real-world' applications of linear systems and, more generally, on methods of
solving linear systems using computers.

The software we shall employ toward this end styles
\href{https://www.sagemath.org/}{SageMath}. It's a free open-source mathematics
software capable of numeric and symbolic manipulation of objects from various
fields of mathematics, linear algebra included. It can be installed on most
operating systems following the
\href{https://doc.sagemath.org/html/en/installation/index.html}{official guide}.

SageMath is essentially a terminal-based software and out of the box offers no
graphical user interface. Upon launch, the user is greeted by a screen similar
to this one:
\begin{Verbatim}
┌────────────────────────────────────────────────────────────────────┐
│ SageMath version 10.4, Release Date: 2024-07-19                    │
│ Using Python 3.12.8. Type "help()" for help.                       │
└────────────────────────────────────────────────────────────────────┘
sage:
\end{Verbatim}
SageMath is mainly built upon C and Python and is \emph{interpreted}, meaning
every piece of code is immediately run without a need for compilation.

Before we focus on applications of linear systems in fields like
\emph{economics} and \emph{physics}, we need to learn to solve them using
SageMath. By far the simplest way to encode linear systems is using matrix
notation. SageMath features the \texttt{Matrix} class which hosts a plethora of
methods for matrix manipulation we are going to make great use of in time.

\myref{Example}{exam:static-equations} contains the system
\[
 \left(
  \begin{matrix*}[r]
   15 & 40\\
   25 & -50
  \end{matrix*}
  \hspace{1mm}
 \right|
 \left.
  \begin{matrix*}[r]
   100\\
   50
  \end{matrix*}
 \right).
\]
Let us solve it using SageMath. The \texttt{Matrix} class expects a matrix to be
defined as a list of rows which are themselves lists of elements. In addition,
we may specify the number set wherein the elements lie. For example,
\begin{Verbatim}
sage: A = \clb{Matrix(}ZZ, [
....:     [\clr{15}, \clr{40}],
....:     [\clr{25}, \clr{-50}],
....: ]\clb{)}
\end{Verbatim}
creates the matrix
\[
 \left(
  \begin{matrix*}[r]
   15 & 40\\
   25 & -50
  \end{matrix*}
 \right)
\]
with entries in $\Z$, the integers. This has a caveat. When we tell SageMath our
matrix contains entries \emph{exclusively} in $\Z$, it will fulfil our wish with
utmost conscientiousness. This means that \texttt{A} can never contain anything
but integers. A problem might emerge should we wish to put it into echelon form
for example. Gauss-Jordan elimination of the matrix \texttt{A} would clearly
require subtracting $(25 / 15)$-multiple of row \texttt{I} from row \texttt{II}.
Assuming the entries of \texttt{A} are solely integers, such an operation is not
permitted. The \texttt{Matrix} class has an in-built method for Gauss-Jordan
elimination. Let us try to use it.
\begin{Verbatim}
sage: A.\clb{echelon_form()}
[  \clr{5} \clr{130}]
[  \clr{0} \clr{350}]
\end{Verbatim}
The result is somewhat unexpected. Thankfully or unfortunately, SageMath is
clever enough to know that simply following the algorithm of the Gauss-Jordan
elimination does not yield an integer matrix. So, it instead follows the
algorithm and then multiplies the matrix by the least common multiple of the
denominators of all entries in order to yield an integer matrix. Beware however,
that trying to solve linear systems whose solutions are rational with integer
matrices might result in an error. To stay in the clear, we instead use the real
numbers throughout the calculation. Not specifying the number set would lead to
SageMath `guessing' it based on the values of the entries -- which are all
integral.

We thus rewrite our matrix \texttt{A} like this:
\begin{Verbatim}
sage: A = Matrix(\clb{RR}, [
....:     [\clr{15}, \clr{40}],
....:     [\clr{25}, \clr{-50}],
....: ])
\end{Verbatim}
We will also create a \texttt{vector} (a \texttt{Matrix} with a single row
basically) of the right hand side of the studied system.
\begin{Verbatim}
sage: b = \clb{vector(}RR, [\clr{100}, \clr{50}]\clb{)}
\end{Verbatim}
The \texttt{Matrix} method for solving a system with a given \texttt{vector} of
right hand side is called \texttt{solve\_right}. Using it gives
\begin{Verbatim}
sage: A.\clb{solve_right(}b\clb{)}
(\clr{4.00000000000000}, \clr{1.00000000000000})
\end{Verbatim}
Since we explicitly required SageMath solve the system over the real numbers, it
returned the solution as a pair of decimals rounded based on a default precision
parameter. We would instead prefer to write the solution as $(4,1)$. Should we
wish to record the solution as a pair of fractions or integers instead, we would
need to define \texttt{A} and \texttt{b} over $\Q$.
\begin{Verbatim}
sage: A = Matrix(\clb{QQ}, [
....:     [\clr{15}, \clr{40}],
....:     [\clr{25}, \clr{-50}],
....: ])
sage: b = vector(\clb{QQ}, [\clr{100}, \clr{50}])
sage: A.solve_right(b)
(\clr{4}, \clr{1})
\end{Verbatim}

\input{linear-systems/applications/numerical-stability}
\input{linear-systems/applications/gauss-jordan-elimination}
\input{linear-systems/applications/input-output}
\input{linear-systems/applications/circuits}

